### 2020-01-23 presentation_templates/LaTex_beamer/slides.tex
Add presentation templates

\documentclass[xcolor=x11names,compress,usenames,dvipsnames,mathsans]{beamer}
\usepackage{textpos}
\usepackage[greek,english]{babel}
\usepackage{natbib}
\usepackage{wrapfig,color,dashrule,epsfig,graphicx}
\usepackage{pstool,psfrag}
\newcommand{\calt}{\color{red} } % alert color
\newcommand{\ccom}{\color{blue} } % commentary color

\newcommand{\redtext}[1]{\color{UOEred}{#1}\color{UOEblue}}

\newcommand{\link}[1]{\textbf{\textcolor{PMS3025}{\underline{#1}}}}
\definecolor{NewRed}{rgb}{0.8, 0.0, 0.0}

 %% Beamer Layout %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usetheme{Edinburgh}
\usecolortheme{edinburgh}
\usefonttheme[onlylarge]{structurebold}
\setbeamertemplate{itemize items}[circle]

\setbeamerfont{title}{size={\fontsize{13}{20}}}
\setbeamercolor*{palette tertiary}{fg=black,bg=black!10} 
\setbeamercolor*{palette quaternary}{fg=black,bg=black!10} 
\setbeamercolor{alerted text}{fg=red}
\setbeamercolor{sidebar}{bg=red!25!blue,fg=white}

\def\ci{\perp\!\!\!\perp}

\setbeamercovered{transparent}

\addtobeamertemplate{footline}{}
{
  \begin{textblock*}{100mm}(.845\textwidth,-0.57cm)

    \includegraphics[height=0.5cm, width=1.8cm]{225px-Epcc_logo.jpg}
  \end{textblock*}
}

\addtobeamertemplate{frametitle}{}
{
  \begin{textblock*}{100mm}(.935\textwidth,-0.645cm) 
	\includegraphics[scale=.1]{crest_rb.eps}
  \end{textblock*}
}

\usepackage{graphicx} % Allows including images
\usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule in tables

%\setbeamercovered{transparent}

%----------------------------------------------------------------------------------------
%	STYLE STUFF -- section title slides
%----------------------------------------------------------------------------------------

\AtBeginSection[]{
  \begin{frame}
  \vfill
  \centering
  \begin{beamercolorbox}[sep=8pt,center,shadow=true,rounded=true]{title}
    \usebeamerfont{title}\insertsectionhead\par%
  \end{beamercolorbox}
  \vfill
  \end{frame}
}

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\title[GitLab demo]{Project Preparation \\ GitLab} 
% The [Short Title] appears at the bottom of every slide, the {Full Title} is only on the title page

\author[author]{Full Name \\
\textit{email@ed.ac.uk} \\
\vspace{1em}
\includegraphics[scale=.8]{logo_colour.pdf}
} % Your name

\date{\today}

\begin{document}

\begin{frame}
\titlepage % Print the title page as the first slide

\end{frame}

\begin{frame}
\frametitle{Overview} % Table of contents slide, comment this block out to remove it
\tableofcontents % Throughout your presentation, if you choose to use \section{} and \subsection{} commands, these will automatically be printed on this slide as an overview of your presentation
\end{frame}

%----------------------------------------------------------------------------------------
%	PRESENTATION SLIDES
%----------------------------------------------------------------------------------------

%------------------------------------------------
%\section{Section Title} % Sections can be created in order to organize your presentation into discrete blocks, all sections and subsections are automatically printed in the table of contents as an overview of the talk
%------------------------------------------------

\begin{frame}
\frametitle{Purpose}
\vfill
GitLab is a version control service provided by the University of Edinburgh. \\
\vfill
\pause
It's purpose in the MSc project work is to:
\begin{itemize}
	\item maintain traceable history
	\item document your progress
	\item share information with your supervisor(s) and markers
\end{itemize}
\pause
\vfill
(and it will be marked accordingly.)
\vfill
\end{frame}

%------------------------------------------------
\begin{frame}
\frametitle{Content}

%Proof of your engagement with the project work and ability to organise your work and materials.
%\pause

Project Information:
	\begin{itemize}
		\item Project Title
		\item Student's Name and UUN
		\item EPCC Supervisor's Name
		\item (External Supervisor's Name, Email and Organisation)
	\end{itemize}

\vfill
\pause

Project Materials:
	\begin{itemize}
		\item initial and final project proposal
		\item meeting minutes
		\item Project Preparation report + slides
		\item Dissertation report + slides
		\item code/scripts
	\end{itemize}

\end{frame}

%------------------------------------------------

\begin{frame}
\frametitle{Further Information}

\begin{itemize}
	\item \href{https://git.ecdf.ed.ac.uk/}{\link{Log in page}}
	\item \href{https://www.wiki.ed.ac.uk/pages/viewpage.action?spaceKey=ResearchServices\&title=Version+Control+Service+-+GitLab}{\link{GitLab wiki}}
\end{itemize}

\end{frame}

%------------------------------------------------

\end{document}